---
title: "Network Analysis Using R"
author: "Martin Schweinberger"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  bookdown::html_document2: default
bibliography: bibliography.bib
link-citations: yes
---

# Introduction

This document is the report for the *Network Analysis Using R* workshop at QUT, Brisbane, 20121212.

Networks, or graphs, consist of nodes (typically represented as dots) and edges (typically represented as lines) and they can be directed or undirected networks. In directed networks, the direction of edges is captured. For instance, the exports of countries. In such cases the lines are directed and typically have arrows to indicate direction. The thinkness of lines can also be ustilized to encode information such as frequency of contact.

Networks can have different layouts.

The *centrality of networks* is measured as 
* Node degree
* Node closeness (the closer two nodes are, the closer connected ethy are)
* Node betweeness
* Edge betweenness

There are two ways to capture the basic structure of a network
* Adjacency matrix (Matrix of 0s and 1s depending on whether the nodes are connected or not)
*Edge list (Two column matrix to indicate which nodes are connected)

```{r nta_intro_01, echo=F, eval = T, message=FALSE, warning=FALSE}
edgelist <- matrix(c(A, A, A, A, A, ), ncol = 2)
```


## Example 

Who is connected with whom on Twitter and who is the most cetral node in the network?

Outline of the workshop
* Install R and R Studio
* Import data from Australian Twittershere
* Tidy data



We will now begin with the analysis. In a first step, the session is prepared by clearing the workspace, setting options, activating packages and functions, as well as loading relevant functions.

```{r nta_01, echo=T, eval = T, message=FALSE, warning=FALSE}
# activate package
library(igraph)
```




